### 2015-02-22 fundamentals/impl.tex
A variety of separate changes: * some typos, repeated phrases and grammar glitches * clarfiy inheritance for institutions wrt deploy/decommission * change framing of non-RIF operation - suggest footnote * strengthen language about the need for the DRE:     basically uncertainty in every angle: archetype defintion, agent entry/exit, resource quality * improve language connecting mathematical constraints to real-world constraints * introduce role of facilities in preference setting * small clarifications for integration and regression tests

A modular, \acrfull{ABM} approach is ideal for solving complex, coupled,
physics-dependent supply chain problems involving material routing, facility
deployment, and regional and institutional hierarchies. Additionally, the choice to
build \Cyclus on open source libraries in modern programming languages enables
both remote and multiprocess execution on a number of platforms. This section begins by
describing the general design features that make \Cyclus both flexible and powerful:
cluster-ready software and dynamic libraries.  Next, the framework has an
\gls{ABM} paradigm, which provides an expansion in performance beyond typical system dynamics
capabilities, is described, focusing on agent interchangeability and the
implementation of the region-institution-facility hierarchy.
Next, the discrete treatment of objects is discussed in terms of both
resources and materials and time via the \gls{DRE}.
Support for users and developers via the \Cycamore library of
archetypes and the experimental toolkit are also presented.
Lastly, the methods for quality assurance are outlined.

@@

Two primary consequences arise from this structure. First, institutions are
nominally
responsible for deploying and decommissioning facilities. Accordingly, advanced
logic regarding building and decommissioning can be obtained by inheriting
those behaviors from the
\Class{Institution} interface. Second, the \Class{Facility} class implements the
\Class{Trader} interface, and thus institutions and regions, respectively, can
adjust the resource flow preferences of their managed facilities. Importantly,
this capability allows for the modeling of preferential regional trading
of resources (e.g., tariffs) as well as preferential institutional trading
(e.g., long-term contracts).

Although concept of the \gls{RIF} hierarchy is provided explicitly, the \Cyclus kernel is
designed to handle hierarchies of arbitrary depth.  It is capable of
running a simulation without regions and institutions (i.e.  only facilities)
or a simulation having archetypes that are the parents of regions.  By
convention, institutions are mainly responsible for facility deployment, but
all archetypes have access to the agent deployment mechanism. \Cyclus enables
the developer to to implement agents that deploy copies of themselves, or that
schedule their own decommissioning.  Because regions, institutions, and
facilities are all just agents, they can all be implemented to exercise the
full set of features offered by the \Cyclus kernel.

@@

One such use case seeks to capture system vulnerability to nefarious
material diversion. Provenance and trade-history of distinct materials is the fundamental
information unit in such studies, and so this type of analysis requires
 discrete simulation of a
target facility and the individual materials modified within it.
Provenance and trade-history of distinct materials is the fundamental
information unit in such studies.
Material risk analysis, therefore, demands that both facilities and materials
should be discretely modeled objects like those in \Cyclus.

@@

\subsubsection{Dynamic Resource Exchange}

Since the \Cyclus simulation paradigm involves both discrete agents and
discrete resources, the logic engine defining resource interaction mechanisms
among agents is crucial. The \gls{DRE} is that critical logic engine on which \Cyclus
simulations are built.
As the \Cyclus supply-demand resolution framework,
the DRE must be able to support arbitrary facilities. Supporting the general
\Cyclus philosophy, facilities are treated as black boxes and a clear
supply-demand communication framework is defined.

@@

Suppliers then respond to the series of requests with a \textit{bid}. A bid
supplies a notion of the quantity and quality of a resource to match a
request. Suppliers may add an arbitrary number of constraints to accompany
bids. For example, an enriched \gls{UOX} supplier may supply both a natural uranium
and SWU constraint. Any potential resource transfer may be denoted as
\textit{exclusive}. An exclusive transfer must either be met fully or not at all
in order to support concepts such as the trading of individual reactor
assemblies.

Finally, requesting facilities, institutions and regions may apply
\textit{preferences} to each potential request-bid pairing based on the proposed
resource transfer. Therefore, an institution can apply a higher preference to a
partner to which it is congenial; similarly, a region may negate any transfers
of material which have a higher uranium enrichment than is allowable.

Given a full definition of supply and demand, the \gls{DRE} may be solved either
optimally using a mathematical program or approximately by a simulation-based
heuristic \cite{gidden_agent-based_2014}. If any trade is denoted as exclusive, e.g., if an
analyst desires an assembly-fidelity model, then either a heuristic must be used
or exchanges must be represented as a \gls{MILP}. If no exclusive trades exist,
a \gls{LP} model may be used. In practice, LPs solve much faster than
\gls{MILP}s. However, both capabilities exist in \Cyclus in order to
provide users with the desired level of fidelity.

@@

The \gls{DRE} is a novel simulation concept in the nuclear fuel cycle domain. It
provides a flexible supply-demand infrastructure, supporting dynamic flows of
resources between agents. Trading between agents can be affected by both the
proposed quality of a resource and agent relationships through the use of
preferences. Accordingly, a wide range of possible effects can be
modeled, from capacitaty-limited fuel supply to international trade agreements.

@@

The current \Cycamore release provide basic functionality, As illustrated in
Figure \ref{fig:simplesim}, the current \Cycamore release provides basic
functionality enabling simple fuel cycle analyses. As future contributions are
vetted, the capabilities in \Cycamore may grow.

@@

In addition to the core functionality of the \Cyclus kernel, which is focused on
the set of capabilities needed to implement an agent-based simulation
with \gls{DRE}, a toolkit is provided to assist developers
and users with related simulation and nuclear engineering tasks. The toolkit is
an actively developed part of \Cyclus, with a primarily forward-looking
focus on supporting interesting \textit{in situ} metric analysis tools.

@@

A series of utility classes are provided to support demand-constrained agent
facility deployment. For example, symbolic function representations of linear,
exponential, and piecewise functions are supported via the
\Class{SymbFunctionFactory} class. Such functions are used with other toolkit
classes to determine commodity demand (e.g., power demand) from user input. Four
mix-in classes provide the basis for in-simulation deployment determination:
\Class{CommodityProducer}, \Class{CommodityProducerManager}, \Class{Builder},
\Class{BuildingManager}. The \Class{CommodityProducer} class provides an
interface for querying the \textit{prototypes} which have the
capacity to produce a given commdity. The \Class{CommodityProducerManager}
provides an interface for registering \Class{CommodityProducer}s and querying
the current capacity (supply) of a commodity. The \Class{Builder} class provides
an interface for querying which prototypes can be built and interacts with the
\Class{BuildingManager}, which orders prototypes to be built. The
\Class{BuildingManager} uses a simple minimum cost algorithm to determine how
many of each prototype, $y_i$, to build given a demand ($\Phi$), capacities
($\phi_i$), and costs ($c_i$).

@@

Simulation science - like experimental science - must distinguish trustworthy
results from untrustworthy ones.
Charles Babbage famously articulated this, ``On two occasions I have been asked,
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right
answers come out?' ... I am not able rightly to apprehend the kind of confusion
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}.
A simulator must assure correctness and guard against the \emph{garbage
in, garbage out} phenomenon.

Multiple strategies, collectively known as \emph{\gls{QA}}, have
been invented to mitigate structural and algorithmic errors
in software. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}}
\cite{sacks_design_1989}, testing, and others.

Nuclear engineering software quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
\Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or by the PyNE toolkit
\cite{biondo_quality_2014}.

Since \Cyclus modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

Verification may be defined as the question, ``Is \Cyclus being built correctly?''
To answer this question, \Cyclus relies on software development best practices
such as testing,
documentation, version control, style guidelines, and continuous integration to
ensure reliability and reproducibility. These verification techniques are
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question,
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway
\cite{huff_extensions_2014}. However, such
exercises are more likely to bring into relief the differences between the modeling
paradigms than supply substantial \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine quality. \gls{UQ} is a process that applies
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle
simulator, cannot make reliable statements about the uncertainty of metrics for a particular fuel
cycle. Any \gls{UQ} results are applicable only to the scenario that is under
analysis by the user. \Cyclus is not a substitute for due diligence on behalf of the fuel
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development
and in the domain of the expert user.

@@

\Cyclus uses the Google Test framework \cite{inc_googletest_2008} as a harness for running unit
tests. Sufficient unit tests are required for any proposed change to the \Cyclus
code base. Currently, \Cyclus implements over 450 unit tests and \Cycamore implements
85.  These cover approximately 65\% of their respective code bases, and these numbers are expected to grow over time.

@@

Integration tests combine multiple elements of the
\Cyclus interface and test that they work correctly with each other.  By analogy,
simply because the gears (units) are made correctly does not imply that the
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of
time. A set of standard input files have thus been constructed.
Python is used to run the same set of \Cyclus simulations as a subprocess for each integration test.  The results are then inspected and compared via Nose \cite{pellerin_nose_2007}, a Python test framework.
In this way, \Cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that
major \Cyclus components are functioning as expected.

@@

Regression tests ensure unexpected backwards-incompatible changes do not
occur over the course of \Cyclus development. Such a change is called a
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute full \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against
the output of the same input file when run with a previous version of \Cyclus,
typically the last released version.
In some sense, regression tests are `dumb' in that they do
not care about the contents of a simulation being correct, only whether or not
it changed. They pick up glaring time-sensitive mistakes that may
not appear elsewhere, though they do not purport to make claims about the
quality of the physics modeled. Regression tests are primarily implemented
in \Cycamore because its archetypes are sophisticated enough
to make interesting simulations.

>>>>

A modular, \acrfull{ABM} approach is ideal for solving complex, coupled,
physics-dependent supply chain problems involving material routing, facility
deployment, and regional and institutional hierarchies. Additionally, the choice to
build \Cyclus on open source libraries in modern programming languages enables
both remote and multiprocess execution on a number of platforms. This section begins by
describing the general design features that make \Cyclus both flexible and powerful:
cluster-ready software and dynamic libraries.  Next, the framework's
\gls{ABM} paradigm, which provides an expansion in performance beyond typical system dynamics
capabilities, is described, focusing on agent interchangeability and the
implementation of the region-institution-facility hierarchy.
Next, the discrete treatment of objects is discussed in terms of both
resources and materials and time via the \gls{DRE}.
Support for users and developers via the \Cycamore library of
archetypes and the experimental toolkit are also presented.
Lastly, the methods for quality assurance are outlined.

@@

Two primary consequences arise from this structure. First, institutions are
nominally
responsible for deploying and decommissioning facilities. Accordingly, advanced
logic regarding building and decommissioning can be implemented on top of  
those behaviors inherited from the
\Class{Institution} interface. Second, the \Class{Facility} class implements the
\Class{Trader} interface, and thus institutions and regions, respectively, can
adjust the resource flow preferences of their managed facilities. Importantly,
this capability allows for the modeling of preferential regional trading
of resources (e.g., tariffs) as well as preferential institutional trading
(e.g., long-term contracts).

% the paragraph below muddies the concept of the RIF hierarchy
% (perhaps an extended footnote?)
% Either way, I prefer something other than ``provided explicitly''.

Although the \Cyclus kernel is designed around the concept of the \gls{RIF} hierarchy, 
it can handle hierarchies of arbitrary depth.  It is capable of
running a simulation without regions and institutions (i.e.  only facilities)
or a simulation having archetypes that are the parents of regions.  By
convention, institutions are mainly responsible for facility deployment, but
all archetypes have access to the agent deployment mechanism. \Cyclus enables
the developer to to implement agents that deploy copies of themselves, or that
schedule their own decommissioning.  Because regions, institutions, and
facilities are all just agents, they can all be implemented to exercise the
full set of features offered by the \Cyclus kernel.

@@

One such use case seeks to capture system vulnerability to nefarious
material diversion. Provenance and trade-history of distinct materials is the fundamental
information unit in such studies, and so this type of analysis requires
 discrete simulation of a
target facility and the individual materials modified within it.
Material risk analysis, therefore, demands that both facilities and materials
should be discretely modeled objects like those in \Cyclus.

@@

% although this abbreviation is provided by the glossary capability, I think
% the DRE is so important that it should be redefined here.

\subsubsection{Dynamic Resource Exchange (DRE)}

Since the \Cyclus simulation paradigm allows discrete agents, based on archetypes
about which the kernel has no knowledge, to enter 
the simulation at arbitrary times and
trade in discrete resources that are not defined \textit{a priori}, 
the logic engine defining resource interaction mechanisms
among agents is crucial. The \gls{DRE} is that critical logic engine on which \Cyclus
simulations are built.
As the \Cyclus supply-demand resolution framework,
the DRE must be able to support arbitrary facilities. Supporting the general
\Cyclus philosophy, facilities are treated as black boxes and a clear
supply-demand communication framework is defined.

@@

Suppliers then respond to the series of requests with a \textit{bid}. A bid
supplies a notion of the quantity and quality of a resource to match a
request. Suppliers may add an arbitrary number of constraints to accompany
bids. For example, an enriched \gls{UOX} supplier may be constrained by its current
inventory of natural uranium or its total capacity to provide enrichment (SWUs), and
attach such constraints to its bids. Any potential resource transfer may be denoted as
\textit{exclusive}. An exclusive transfer must either be met fully or not at all
in order to support concepts such as the trading of individual reactor
assemblies.

Finally, requesting facilities, institutions and regions may apply
\textit{preferences} to each potential request-bid pairing based on the proposed
resource transfer. Facilities can apply arbitrary complex logic to rank the bids 
that they have received, whether based on the quantity available in each bid or
on the quality of each bid, and the consequent implications of the physics behavior
of that facility. In addition, an institution can apply a higher preference to a
partner to which it is congenial; similarly, a region may negate any transfers
of material which have a higher uranium enrichment than is allowable.

Given a full definition of supply and demand, the \gls{DRE} may be solved either
optimally using a mathematical program or approximately by a simulation-based
heuristic \cite{Gidden2014}. If any trade is denoted as exclusive, e.g., if an
analyst desires an assembly-fidelity model, then either a heuristic must be used
or exchanges must be represented as a \gls{MILP}. If no exclusive trades exist,
a \gls{LP} model may be used. In practice, LPs solve much faster than
\gls{MILP}s. However, both capabilities exist in \Cyclus in order to
provide users with the desired level of fidelity.

@@

The \gls{DRE} is a novel simulation concept in the nuclear fuel cycle domain. It
provides a flexible supply-demand infrastructure, supporting dynamic flows of
resources between agents, even as those agents enter and leave the simulation, and 
even when those agents are defined by archetypes of arbitrary complexity. Trading 
between agents can be affected by both the
proposed quality of a resource and agent relationships through the use of
preferences. Accordingly, a wide range of possible effects can be
modeled, from capacitaty-limited fuel supply to international trade agreements.

@@

As illustrated in
Figure \ref{fig:simplesim}, the current \Cycamore release provides basic
functionality enabling simple fuel cycle analyses  As future contributions are
vetted, the capabilities in \Cycamore may grow.

@@

In addition to the core functionality of the \Cyclus kernel, which is focused on
the set of capabilities needed to implement an agent-based simulation
with \gls{DRE}, a toolkit is provided that assists developers
and users with related simulation and nuclear engineering tasks. The toolkit is
an actively developed part of \Cyclus, with a primarily forward-looking
focus on supporting interesting \textit{in situ} metric analysis tools.

@@

A series of utility classes are provided to support demand-constrained agent
facility deployment. Symbolic function representations of linear,
exponential, and piecewise functions are supported via the
\Class{SymbFunctionFactory} class. Such functions are used with other toolkit
classes to determine commodity demand (e.g., power demand) from user input. Four
mix-in classes provide the basis for in-simulation deployment determination:
\Class{CommodityProducer}, \Class{CommodityProducerManager}, \Class{Builder},
\Class{BuildingManager}. The \Class{CommodityProducer} class provides an
interface for querying the \textit{prototypes} which have the
capacity to produce a given commdity. The \Class{CommodityProducerManager}
provides an interface for registering \Class{CommodityProducer}s and querying
the current capacity (supply) of a commodity. The \Class{Builder} class provides
an interface for querying which prototypes can be built and interacts with the
\Class{BuildingManager}, which orders prototypes to be built. The
\Class{BuildingManager} uses a simple minimum cost algorithm to determine how
many of each prototype, $y_i$, to build given a demand ($\Phi$), capacities
($\phi_i$), and costs ($c_i$).

@@

Simulation science - like experimental science - has since the beginning had the
problem of identifying trustworthy results from nonsensical results or noise.
Charles Babbage famously articulated this, ``On two occasions I have been asked,
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right
answers come out?' ... I am not able rightly to apprehend the kind of confusion
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}.
A simulator must both guard against the \emph{garbage in, garbage out} phenomenon, and also assure correctness.

Multiple strategies, collectively known as \emph{\gls{QA}}, have
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}}
\cite{sacks_design_1989}, testing, and others. \gls{VV} and \gls{UQ} come from
a distinctly computational science background while testing \emph{et al.} come from software development.

Nuclear engineering software quality is often governed by \gls{NQA1}, an
\gls{ASME} specification
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}.
Although \gls{NQA1} is primarily
used for designing reactors, it is general enough to apply to
\Cyclus. \Cyclus has adopted an \emph{agile} development process
\cite{larman_agile_2004},
interpreting \gls{NQA1} in a manner similar to the process adopted by the
\gls{DOE} within \gls{NEAMS} \cite{neams_nuclear_2013} or the
PyNE toolkit \cite{biondo_quality_2014}.

Quality assurance is an on-going process throughout the
entire life of the software. As a simulator where most modeling decisions are made
by third party archetype developers or users, the most important feature of
\gls{QA} is verification. Validation may be impossible and \gls{UQ} is beyond
the scope of the kernel.

Verification may be defined as the question, ``Is \Cyclus being built correctly?''
To answer this question, we turn to the software development of notions of testing,
documentation, version control, style guidelines, and continuous integration.
This suite of process controls supplies mechanism for reliable and reproducible
software. The impetus to implement these controls correctly is particularly strong in a scientific
context because of the emphasis on reproducibility and provenance. These verification techniques are
discussed individually in the sections that follow.

Validation, in contrast,  may be defined as the question,
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator, longitudinal
validation is not possible. Nonetheless, code-to-code comparisons with fuel cycle
simulators that use other modeling paradigms are underway, if nascent. However, such
exercises are more likely to bring into relief the differences between the modeling
paradigms than supply substantial \gls{QA} and validation.

Lastly, \gls{UQ} is a process that is used on specific simulation
instantiations to statistically determine quality. \gls{UQ} is a process that applies
to the running of many \Cyclus simulations. The kernel itself, as a generic fuel cycle
simulator, cannot make reliable statements about the uncertainty of a particular fuel
cycle. Any \gls{UQ} results are applicable only to the scenario that is under
analysis by the user. \Cyclus is not a panacea for due diligence on behalf of the fuel
cycle modeler. Thus, \gls{UQ} is envisioned to be beyond the scope of core development
and in the domain of the expert user.

@@

\Cyclus uses the Google Test framework \cite{inc_googletest_2008} as a harness for running unit
tests. Sufficient unit tests are required for any proposed change to the \Cyclus
code base. Currently \Cyclus implements over 450 unit tests and \Cycamore implements
85.  These cover approximately 65\% of their respective code bases, and their numbers are expected to grow over time.

@@

Integration tests combine multiple elements of the
\Cyclus interface and test that they work correctly with each other.  By analogy,
simply because the gears (units) are made correctly does not imply that the
clock (integration) will run smoothly, run at all, or give the correct time.
In \Cyclus and \Cycamore, integration tests are performed by running sample
simulations for scenarios that are sufficiently simple to predict the results and verifying 
that results match those predictions. A set of standard input files have thus been constructed.
Python is used to run the same set of \Cyclus simulations as a subprocess for each integration test. 
 The results are then inspected and compared via Nose \cite{pellerin_nose_2007}, a Python test framework.
In this way, \Cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that
major \Cyclus components are functioning as expected.

@@

Regression tests ensure significant unintended changes do not
occur over the course of \Cyclus development. Such a change is called a
\emph{regression} because the new version is almost always wrong.
Regression tests are implemented similarly to integration tests.
Nose is used to execute full \Cyclus simulations whose results
are then compared. In this category, however, the comparison is done against
the output of the same input file when run with a previous version of \Cyclus,
typically the last released version.
In some sense, regression tests are `dumb' in that they do
not care about the contents of a simulation being correct, only whether or not
it changed. They pick up glaring time-sensitive mistakes that may
not appear elsewhere, though they do not purport to make claims about the
quality of the physics modeled. Regression tests are primarily implemented
in \Cycamore because its archetypes are sophisticated enough
to make interesting input files. Only a handful ($<10$) of time steps are needed
for regression testing.
